Modal Abstractions for Virtualizing Memory Addresses

### 2023-11-15 examples.tex
clarification
continuing to edit/compress/clarify, make figure 9 slightly more compact

@@

;; Assume the save-space is in rdi,
;; load-space in rsi
;; No need to save/restore caller-save regs
;; Save yielding context

@@

mov 8[rdi], rsp
mov 16[rdi], rbp
mov 24[rdi], r12
mov 32[rdi], r13
mov 40[rdi], r14

@@

;; This switches to a new stack
;; This stack may not be mapped in the
;; current address space!

@@

mov rbp, 16[rsi]
mov r12, 24[rsi]
mov r13, 32[rsi]
mov r14, 40[rsi]

>>>>

We can see that this specification neatly encapsulates reasoning about invariants on the page tables in terms of general modal assumptions.\todo{speculative
but ought to turn out this way}

@@

;; Assume the save-space is in rdi, load-space in rsi
;; First, save yielding context

@@

... ;; mov rsp, rbp, r12, r13, r14, saved to offsets 8, 16, 24, 32, and 40 from rdi

@@

;; This switches to a new stack, which may not be mapped in the current address space!

@@

... ;; mov rbp, r12, r13, r14, to offsets 16, 24, 32, and 40 from rsi

### 2023-11-15 programlogic.tex
continuing to edit/compress/clarify, make figure 9 slightly more compact

\lstinline|L4_L1_PointsTo| works by
chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
each table level, and using the physical page address embedded in each entry.
For example, 
  the first level address translation to get the L4 entry (\lstinline|l4e|) 
  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the 
  physical address of the start of the L4 table,
  and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
  Then at that physical location is the appropriate entry in the L4 table \textsf{l4e}.
  Subsequent levels of the page table walk work similarly.
The statement of these assertions is simplified by the use of our split physical points-to assertions, since
each level of tables is page-sized.\footnote{We do not address superpages and hugepages in this paper.}
This helper definition is also a more explicit \textsf{vProp}, explicitly binding a value for \lstinline|cr3| and using it
to start the translation process.

@@

\todo[inline,color=cyan]{Colin edit frontier}

\subsection{Aliasing/Sharing Physical Pages}
  \label{sec:sharingpages}  
  One might have already observed that the virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
  is too strong to specify some operations a virtual memory manager may need to do, such as
  move one level of the page table to a different physical location while preserving all virtual-to-physical mappings.\footnote{
    x86-64 hardware, like other architectures, includes a feature (which we do not formalize assertions for) to
    replace an L1 page table address in an L2 entry with a pointer to a \emph{larger} 2MB page (called super-pages), 
    or replace an L2 page table address in an L3 entry with a pointer to a 1GB page (called huge-pages).
  }
   To speak more concretely, the fractional permissions on the page-table-walk assertion ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
  ensure fractional ownership of the four physical page table entries accessed by the hardware during address translation,
  as well as the physical memory holding the data of the virtual points-to.
  However, this formulation stores knowledge of the page table walk's details with ownership of the backing physical
  memory.
  Updating any of these mappings (e.g., moving the page tables in physical memory, as in coallescing for
  superpages or hugepages)
  would require explicitly collecting all virtual points-to facts that traverse affected entries.
  It is preferable to permit the page tables themselves to be updated independently of the virtual points-to assertions,
  so long as those updates preserve the same virtual-to-physical translations.
\newcommand{\vale}{\textsf{val}}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\;\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} \\
    \quad
    \exists \paddr\ldotp
    \exists \delta\ldotp
    \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{\begin{tabular}{c}Find ghost map for this\\address space\end{tabular}} \ast 
  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{\begin{tabular}{c}Fractional ownership of \\virt-to-phys translation\end{tabular}}\ast 
  \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{q}\}\; \vale}_\text{\begin{tabular}{c}Value at\\physical location\end{tabular}}
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}

  Intuitively, the definition in Figure \ref{fig:strongvirtualpointsto} is too strong because the virtual points-to
  assertion there tracks too much information: when writing programs that access memory via virtual addresses,
  most code does not care \emph{which physical memory locations are involved in address translation}: it only cares
  that virtual address translation would succeed. The necessary information about the physical page table walk
  must still be tracked, but can be tracked separately from the virtual points-to assertion itself.

  We take the approach of separating the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) 
  from the virtual pointsto relation. 
  To do so, we take advantage of an Iris construction which permits tracking a map in ghost state that maps
  virtual addresses to the physical addresses they translate to, and has for each key in the map (i.e.,
  each virtual address) a token. Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
  requires full ownership of the token. In addition, we keep an invariant that for each $va\mapsto pa$ mapping
  in this map, there are \emph{physical} resources sufficient to ensure that address translation for $va$
  will resolve in the hardware to $pa$. But these physical resources, which specify \emph{which} physical locations
  are involved in the page table walk, are now separated from but consistent with
  the knowledge that such resources exist (which is embodied by the token for $va$, which tracks that $va$ maps to $pa$
  in the ghost map). Thus we can store the token which summarizes the translation and ensures it exists in the virtual
  points-to, and keep the ghost map and the invariant that every mapping in the ghost map has corresponding physical resources
  for translation in a separate global invariant for each address space.

  % To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$. 
  % We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
  In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
  map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
  The resource algebra itself is instantiated as:
  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
We replace the physical
 $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$
with
fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
 whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)

We place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in a per-address-space invariant
($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership, when combined with the
fraction stored in the virtual points-to, permits changes in the virtual-to-physical translations.
By itself, it allows changes to the page tables as long as they do not change the overall mappings from virtual
to physical addresses.

@@

Figure \ref{fig:peraspaceinvariant} describes the resources for each address space, which are collected
for the full address space, but separated from virtual points-to assertions.
$\Theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
for the \emph{current} address space, corresponding to the currently installed page tables (the ones
indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially-quantified $\delta$
is the ghost name for the authoritative copy of this map (again, for the current address space).
$m$ is the (full) logical map from the various page table bases to ghost names for the 
per-address-space mappings like $\Theta$.
It is the full version of the map named by the ghost name $\delta{}s$ in
Figure \ref{fig:virtualpointstosharing}; there is only one such map in the system.
This is why Figure \ref{fig:virtualpointstosharing} asserts there is some ghost name in that (global) map
for the current address space.
The global map itself is tracked by a resource algebra 
\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \textsc{GhostName},\mathord{=}) )) \]

$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space:
It ensures that the current address space is valid (it has an entry in $m$, whose corresponding
authoritative map matches $\Theta$), and asserts ownership of the appropriate fractional ownership
of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).

@@

\paragraph{From A Single Address Space to Many}
In principle the resources from each individual address space should be collected into a single
shared invariant, and for each memory access,
a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
for the current address space should be extracted from this global resource, used to prove correctness of the
memory access, and put back.
In this paper we focus on the middle section, explicity identifying resources for each
address space. The reason for this is that in practice, this global resource is also
the place where kernel-specific assumptions, such as guaranteeing that certain virtual address range
was mapped in all address spaces, would be enforced. This paper focuses on the
reasoning principles behind the virtual memory access, and we leave the use of this within a larger
kernel to future work.

\subsection{Address-Space Management}
\label{sec:aspacemanagement}
So far, we have introduced logical abstractions for a single address space. However, virtual memory management
implementations typically handle more than one address-space,
and doing this requires assertions which talk about other address spaces, and means to switch address spaces.
% requiring a means to  which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space

@@

% In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition. To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the  \lstinline|cr3|, it is expected to mention resources of two different address-spaces in the specification. In the precondition,  \lstinline|cr3| is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}. The other address-space with the root address R1 has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0. This pattern of specification requires resources being mentioned under the \textit{world} they are valid, i.e. a modal-framing within the specification. Once the current address-space becomes the one with root address R1, i.e.  \lstinline|cr3| is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
\begin{figure}
\[
  \begin{array}{r@{\qquad}l}
    \hbox{(\TirNameStyle{ModalAddressSpace})} &
         [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} & (\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))\\
    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} & [r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})\\
    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} &   [r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})\\    
    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} &    [r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})\\   
    % \hbox{(\TirNameStyle{ModalAddressSpacePure})} &  \textsf{pure\_vprop P}(\textsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \textsf{P}\\
    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} &  \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
    % \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
    % \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} 
    &\textsf{Fact P} \vdash  ([r](\textsf{P}) \dashv\vdash (\textsf{P}) %(\textsf{vProp }\Sigma))\\
  \end{array}
  \]
  \caption{Other-space Modality and Its Laws}
  \label{fig:modaldef}
  \end{figure}
Figure \ref{fig:modaldef} gives the definition of our modal operator for asserting truth of a modal
(address-space-contingent) assertion \emph{in another address space}, which we call
the \emph{other-space} modality. The definition itself is not
particularly surprising --- as our modal assertions are semantically predicates on a page table root (physical)
address, the assertion $[r](P)$ is a modal assertion which ignores the (implicit) current page table root,
and evaluates the truth of $P$ as if $r$ were the page table root. This simple definition is 
what we need in order for assertions in one address space to talk about assertions in another.

>>>>

\lstinline|L4_L1_PointsTo| works by
chaining together the entries for each level, using the sequence of table offsets from the address being translated to index
each table level, and using the physical page address embedded in each entry.\footnote{
  The fractions \lstinline|q1| through \lstinline|q4| represent the fractional ownership of each entry based on how many
  word-aligned addresses might need to share the entry.
}
For example, 
  the first level address translation to get the L4 entry (\lstinline|l4e|) 
  uses the masks \textsf{l4M52} with the current \lstinline|cr3| to get the 
  physical address of the start of the L4 table,
  and \textsf{l4off} with the virtual address being translated to compute the correct offset within that table.
  Then at that physical location is the appropriate entry in the L4 table \textsf{l4e}.
  Subsequent levels of the page table walk work similarly.
The statement of these assertions is simplified by the use of our split physical points-to assertions, since
each level of tables is page-sized.\footnote{We do not address superpages and hugepages in this paper.}
This helper definition is also a more explicit \textsf{vProp}, explicitly binding a value for \lstinline|cr3| and using it
to start the translation process.\todo{Ismail: aren't phys ptstos vProps? Or iProps?}

@@

\subsection{Aliasing/Sharing Physical Pages}
  \label{sec:sharingpages}  
  The virtual points-to definition shown in Figure \ref{fig:strongvirtualpointsto} 
  is too strong to specify some operations a virtual memory manager may need to do, such as
  move one level of the page table to a different physical location while preserving all virtual-to-physical mappings.\footnote{
    x86-64 hardware, like other architectures, includes a feature (which we do not formalize assertions for) to
    replace an L1 page table address in an L2 entry with a pointer to a \emph{larger} 2MB page (called super-pages), 
    or replace an L2 page table address in an L3 entry with a pointer to a 1GB page (called huge-pages).
  }
  The use of $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}'s
  virtual points-to definition'
   stores knowledge of the page table walk's details with ownership of the backing physical
  memory.
  Updating any of these mappings (e.g., moving the page tables in physical memory, as in coallescing for
  superpages or hugepages)
  would require explicitly collecting all virtual points-to facts that traverse affected entries.
  It is preferable to permit the page tables themselves to be updated independently of the virtual points-to assertions,
  so long as those updates preserve the same virtual-to-physical translations.
\newcommand{\vale}{\textsf{val}}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\;\vale : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} \\
    \quad
    \exists \paddr\ldotp
    \exists \delta\ldotp
    \underbrace{(\lambda\mathit{cr3val}\ldotp\sumapaces{\mathit{cr3val}}\delta)}_\text{Find addr.\;space invariant} \ast 
  \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation}\ast 
  \underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{q}\}\; \vale}_\text{Physical location}
\end{array}
\]
\vspace{-1em}
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}

  Intuitively, the definition in Figure \ref{fig:strongvirtualpointsto} is too strong because the virtual points-to
  assertion there tracks too much information: when writing programs that access memory via virtual addresses,
  most code does not care \emph{which physical memory locations are involved in address translation}: it only cares
  that virtual address translation would succeed. The necessary information about the physical page table walk
  must still be tracked, but can be tracked separately from the virtual points-to assertion itself.
  In practice the decisions about which virtual addresses are valid rest not with code posessing a virtual address, but with
  the virtual memory manager --- and its invariants.

  We take separate the physical page-table-walk  
  from the virtual pointsto relation, replacing it with ghost state that merely guarantees the address translation would succeed.
  Iris includes a construction which permits tracking a ghost map 
  from virtual addresses to the physical addresses they translate to, and has for each key in the map (i.e.,
  each virtual address) a token. Updating the virtual-to-physical (ghost) mapping for a virtual address $va$
  requires full ownership of the token. Then the \emph{virtual memory manager's invariant} ensures that for each $va\mapsto pa$ mapping
  in this map, there are \emph{physical} resources sufficient to ensure that address translation for $va$
  will resolve in the hardware to $pa$ --- $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$.
  This allows modification of the page tables themselves as long as the changes do not invalidate overall virtual-to-physical
  translation.
  Consistency follows from replacing the page table walk with ownership of the ghost map token for the relevant ghost map
  entry, preventing the overall translation from being changed separately from the virtual points-to.

  %Thus the specification of \emph{which} physical addresses support translation is separated from the virtual points-to.
  % But these physical resources, which specify \emph{which} physical locations
  %are involved in the page table walk, are now separated from but consistent with
  %the knowledge that such resources exist (which is embodied by the token for $va$, which tracks that $va$ maps to $pa$
  %in the ghost map). Thus we can store the token which summarizes the translation and ensures it exists in the virtual
  %points-to, and keep the ghost map and the invariant that every mapping in the ghost map has corresponding physical resources
  %for translation in a separate global invariant for each address space.

  % To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$. 
  % We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
  In Iris this is realized by using an authoritative resource algebra: there is a single \emph{authoritative} global copy of the (ghost)
  map caching virtual-to-physical address translations, and for each entry a read only \emph{partial} ownership of that key-value pair.
  The resource algebra itself is instantiated as:
  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
We replace the physical
 $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in the virtual points-to
with
fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$),
guaranteeing that ghost map contains an mapping from the virtual address ($\vaddr$) to a physical address ($\paddr$).

We place the authorative ownership of the ghost translation $\mathcal{A}\textsf{PTableWalk}$ in a per-address-space invariant
$\mathcal{I}$\textsf{ASpace} (Figure \ref{fig:peraspaceinvariant}). 
$\mathcal{I}$\textsf{ASpace} alone allows changes to the page tables 
that preserve overall virtual-to-physical translations.
When combined with the
fragment stored in the virtual points-to (Figure \ref{fig:virtualpointstosharing}),
changes in the virtual-to-physical translations become possible.

@@

Figure \ref{fig:peraspaceinvariant} describes the resources for each address space.
$\Theta$ is the logical map from virtual addresses to the physical addresses they should translate to, 
for the \emph{current} address space, corresponding to the currently installed page tables (the ones
indicated by \lstinline|cr3|). In Figure \ref{fig:virtualpointstosharing}, the existentially-quantified $\delta$
is the ghost name for the authoritative copy of this map (again, for the current address space).
$m$ is the (full) logical map from the various page table bases to ghost names for the 
per-address-space mappings like $\Theta$.
It is the full version of the map named by the ghost name $\delta{}s$ in
Figure \ref{fig:virtualpointstosharing}; there is only one such map in the system.
This is why Figure \ref{fig:virtualpointstosharing} asserts there is some ghost name in that (global) map
for the current address space.
The global map itself is tracked by a resource algebra 
\[\mathcal{A}\textsf{spaces} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  ( (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \textsc{GhostName},\mathord{=}) )) \]

$\mathcal{I}\textsf{ASpace}(\Theta,m)$ is then the resources for the current address space,
ensuring that the current address space is valid (has an entry in $m$ whose corresponding
authoritative map matches $\Theta$), and asserting ownership of the appropriate fractional ownership
of the physical page table walk for each virtual-to-physical translation in $\Theta$ (for the current address space).

@@

\paragraph{From A Single Address Space to Many}
In principle the resources from each individual address space should be collected into a single
shared invariant, and for each memory access,
a fragment of this corresponding to $\mathcal{I}\textsf{ASpace}(\Theta,m)$
for the current address space should be extracted from this global resource, used to prove correctness of the
memory access, and put back.
In this paper we focus on the middle section, explicity identifying resources for each
address space. The reason for this is that in practice, this global resource is also
the place where kernel-specific assumptions, such as guaranteeing that certain virtual address range
was mapped in all address spaces, would be enforced. This paper focuses on the
reasoning principles behind the virtual memory access, and we leave the use of this within a larger
kernel to future work. We do, however, deal with the existence of multiple address spaces.

\subsection{Address-Space Management}
\label{sec:aspacemanagement}
So far, we have introduced logical abstractions for a single address space, but VMMs
 handle more than one address-space,
and doing so requires a way to talk about other address spaces, and means to switch address spaces.
% requiring a means to  which we abstract as a ghost map from root address of an address-space to a namespace labelling the address space

@@

% In Figure \ref{fig:addrswitch}, we see two different address-spaces rooted at R0 and R1 in the pre/post condition. To be put into words, while switching an address-space by executing \lstinline|mov_ctl|) updating the  \lstinline|cr3|, it is expected to mention resources of two different address-spaces in the specification. In the precondition,  \lstinline|cr3| is loaded with R0 and all the virtual-pointsto assertions $a_0 \mapsto_{\textsf{v}}p_0$ and $a_1 \mapsto_{\textsf{v}}p_1$ are in the current-view of the memory and valid in the ambient logic \textsf{vProp}. The other address-space with the root address R1 has its own resources ($b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$) which are invalid under the current address space rooted at R0. This pattern of specification requires resources being mentioned under the \textit{world} they are valid, i.e. a modal-framing within the specification. Once the current address-space becomes the one with root address R1, i.e.  \lstinline|cr3| is loaded with R1, the context-resources $b_0 \mapsto_{\textsf{v}}p_3$ and $b_1 \mapsto_{\textsf{v}}p_4$ are loaded into the current-view of the memory, and all the other resources that are valid under R0 are introduced into the modal context rooted at R0
\begin{figure}
\[
  \begin{array}{r@{\qquad}l}
    \hbox{(\TirNameStyle{ModalAddressSpace})} &
         [r](\mathsf{P}) \;: \mathsf{vProp} \; \Sigma \; \stackrel{\triangle}{=} \; \lambda \_,\; \mathsf{P}~\mathsf{r}.\\
    \hbox{(\TirNameStyle{ModalAddressSpaceMono})} & (\textsf{P} \vdash \textsf{Q}) \vdash  ([r](\textsf{P}) \vdash  [r](\textsf{Q}))\\
    \hbox{(\TirNameStyle{ModalAddressSpaceSep})} & [r](\textsf{P} \ast \textsf{Q}) \dashv\vdash ([r](\textsf{P}) \ast [r](\textsf{Q})\\
    \hbox{(\TirNameStyle{ModalAddressSpaceAnd})} &   [r](\textsf{P}\land \textsf{Q}) \dashv\vdash [r](\textsf{P})\land [r](\textsf{Q})\\    
    \hbox{(\TirNameStyle{ModalAddressSpaceOr})} &    [r](\textsf{P}\lor \textsf{Q}) \dashv\vdash [r](\textsf{P})\lor [r](\textsf{Q})\\   
    % \hbox{(\TirNameStyle{ModalAddressSpacePure})} &  \textsf{pure\_vprop P}(\textsf{iPropI }\Sigma)\textsf{ : vProp }\Sigma \stackrel{\triangle}{=} \; \lambda \_ \ldotp \textsf{P}\\
    \hbox{(\TirNameStyle{ModalAddressSpaceFact})} &  \textsf{Fact P} \stackrel{\triangle}{=} \;\forall \; ,r \; r' \ldotp  \textsf{P r} \dashv\vdash \textsf{P r'}\\
    % \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} \\
    % \qquad\qquad \textsf{Fact pure\_vprop P} \vdash  ([r](\textsf{pure\_vprop P}) \dashv\vdash (\textsf{pure\_vprop P}) (\textsf{vProp }\Sigma))\\
    \hbox{(\TirNameStyle{ModalAddressSpaceFactIntroElim})} 
    &\textsf{Fact P} \vdash  ([r](\textsf{P}) \dashv\vdash (\textsf{P}) %(\textsf{vProp }\Sigma))\\
  \end{array}
  \]
  \caption{Other-space Modality and Its Laws}
  \label{fig:modaldef}
  \end{figure}
Figure \ref{fig:modaldef} gives the definition of our modal operator for asserting truth of a modal
(address-space-contingent) assertion \emph{in another address space}, which we call
the \emph{other-space} modality. The definition itself is not
particularly surprising --- as our modal assertions are semantically predicates on a page table root (physical)
address, the assertion $[r](P)$ is a modal assertion which ignores the (implicit) current page table root,
and evaluates the truth of $P$ as if $r$ were the page table root. This simple definition is 
what we need in order for assertions in one address space to talk about assertions in another.
\looseness=-1

\todo[inline,color=cyan]{Colin revision frontier}
